---
title: "F0_analysis"
author: "Giang Le"
date: "11/5/2021"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---
title: "F0 Analysis"
author: "Giang Le"
date: "11/05/2021"
output:
  pdf_document: 
    latex_engine: xelatex
  html_document:
    df_print: paged
---

## Plan for Analysis
- Plot F0 and 95% confidence intervals on aggregate.
- Plot F0 and 95% confidence intervals by gender, by noise levels, by vowels,
by vowel and noise levels, by environment (isolation or in a carrier sentence) and also create a regression line for each categorization.
- Summarize findings & email Yan
Other notes
Also leave one speaker out and try classification. If the reslt is good we can
use it to predict tone produced on what kind of level (leave the noise level aside).
The noise shouldn't matter much at this stage. (18 levels?)
look at similarity of tones (using clustering). 
- plot F1 and F2 and vowel spaces.



```{r}

# Read in all the F0 contour info. All of these f0.mat files have been downsampled
# in MatLab during linear interpolation (see getf0.mat in the matlab codes)

# matlab code:
# myFolder = '\\client\h$\Desktop\ProsodyPro\m-3-78\channel1';
# myFiles = dir(fullfile(myFolder,'*.wav')); %gets all wav files
# 
# for k = 1:length(myFiles)
#   baseFileName = myFiles(k).name;
#   fullFileName = fullfile(myFolder, baseFileName);
#   fprintf(1, 'Now reading %s\n', fullFileName);
#   [y, Fs] = audioread(fullFileName);
#   [f0, ~] = pitchRocco(y, Fs);
#   i = 1:length(f0);
#   i_new =  linspace(min(i), max(i), 20);
#   f0_downsampled = interp1(i, f0, i_new, 'linear');  
#   fid= fopen(fullfile(myFolder, 'F0.mat'),'a');
#   fprintf(fid, '%s ', baseFileName);
#   fprintf(fid, '%f ', f0_downsampled);
#   fprintf(fid,'\n');
#   fclose(fid);
# end

# Step 1: Read in all F0.mat files into f0Files and assign a name for each

f0Files <- lapply(Sys.glob("*/channel1/F0.mat"), read.table)
length(f0Files)

f0_f_1_78 <- as.data.frame(f0Files[1])
f0_f_1_90 <- as.data.frame(f0Files[2])
f0_f_1_q <- as.data.frame(f0Files[3])
f0_f_2_78 <- as.data.frame(f0Files[4])
f0_f_2_90 <- as.data.frame(f0Files[5])
f0_f_2_q <- as.data.frame(f0Files[6])
f0_m_1_78 <- as.data.frame(f0Files[7])
f0_m_1_90 <- as.data.frame(f0Files[8])
f0_m_1_q <- as.data.frame(f0Files[9])
f0_m_2_78 <- as.data.frame(f0Files[10])
f0_m_2_90 <- as.data.frame(f0Files[11])
f0_m_2_q <- as.data.frame(f0Files[12])
f0_m_3_78 <- as.data.frame(f0Files[13])
f0_m_3_90 <- as.data.frame(f0Files[14])  
f0_m_3_q <- as.data.frame(f0Files[15]) 

```
 
 
```{r}

# Step 2: Add column names for all dataframes.

numbers <- 1:22
cols <- c("sound.name",numbers)
colnames(f0_f_1_78) <- cols
colnames(f0_f_1_90) <- cols
colnames(f0_f_1_q) <- cols
colnames(f0_f_2_78) <- cols
colnames(f0_f_2_90) <- cols
colnames(f0_f_2_q) <- cols
colnames(f0_m_1_78) <- cols
colnames(f0_m_1_90) <- cols
colnames(f0_m_1_q) <- cols
colnames(f0_m_2_78) <- cols
colnames(f0_m_2_90) <- cols
colnames(f0_m_2_q) <- cols
colnames(f0_m_3_78) <- cols
colnames(f0_m_3_90) <- cols
colnames(f0_m_3_q) <- cols


# Assigning noise level
f0_f_1_78$noise = 78
f0_f_1_90$noise = 90
f0_f_1_q$noise = 0

f0_f_2_78$noise = 78
f0_f_2_90$noise = 90
f0_f_2_q$noise = 0

f0_m_1_78$noise = 78
f0_m_1_90$noise = 90
f0_m_1_q$noise = 0

f0_m_2_78$noise = 78
f0_m_2_90$noise = 90
f0_m_2_q$noise = 0

f0_m_3_78$noise = 78
f0_m_3_90$noise = 90
f0_m_3_q$noise = 0


# Assigning gender variable (0 for female and 1 for male)
f0_f_1_78$gender = 0
f0_f_1_90$gender = 0
f0_f_1_q$gender = 0

f0_f_2_78$gender = 0
f0_f_2_90$gender = 0
f0_f_2_q$gender = 0

f0_m_1_78$gender = 1
f0_m_1_90$gender = 1
f0_m_1_q$gender = 1

f0_m_2_78$gender = 1
f0_m_2_90$gender = 1
f0_m_2_q$gender = 1

f0_m_3_78$gender = 1
f0_m_3_90$gender = 1
f0_m_3_q$gender = 1


### Concatenate all dataframes
f0_reports <- rbind(f0_f_1_78, f0_f_1_90, f0_f_1_q,
                    f0_f_2_78, f0_f_2_90, f0_f_2_q,
f0_m_1_78, f0_m_1_90, f0_m_1_q,
f0_m_2_78, f0_m_2_90, f0_m_2_q,
f0_m_3_78, f0_m_3_90, f0_m_3_q)

dim(f0_reports)
```

According to the report, this dataframe has 4,732 records. In total, from 5 speakers,
we have 4732 sound segments to analyze.

```{r}

# Assign tone values.

# Extracting a substring that contains only the syllable names.
nameswithoutwav <- sapply(strsplit(f0_reports[,1], split=".", fixed=TRUE), "[", 1)
f0_reports$syllable.names <- sapply(strsplit(nameswithoutwav, split="_", fixed=TRUE), "[", 1)

f0_reports$tone <- ifelse(grepl("a", f0_reports$syllable.name, ignore.case=T), "A1",
ifelse(grepl("^tát", f0_reports$syllable.name, ignore.case=T), "D1",
ifelse(grepl("^tạt", f0_reports$syllable.name, ignore.case=T), "D2",
ifelse(grepl("^tết", f0_reports$syllable.name, ignore.case=T), "D1",
ifelse(grepl("^tệt", f0_reports$syllable.name, ignore.case=T), "D2",
ifelse(grepl("^tút", f0_reports$syllable.name, ignore.case=T), "D1",
ifelse(grepl("^tụt", f0_reports$syllable.name, ignore.case=T), "D2",
ifelse(grepl("à", f0_reports$syllable.name, ignore.case=T), "A2",
ifelse(grepl("á", f0_reports$syllable.name, ignore.case=T), "B1",
ifelse(grepl("ả", f0_reports$syllable.name, ignore.case=T), "C1",
ifelse(grepl("ã", f0_reports$syllable.name, ignore.case=T), "C2",
ifelse(grepl("ạ", f0_reports$syllable.name, ignore.case=T), "B2",
ifelse(grepl("ê", f0_reports$syllable.name, ignore.case=T), "A1",
ifelse(grepl("ề", f0_reports$syllable.name, ignore.case=T), "A2",
ifelse(grepl("ế", f0_reports$syllable.name, ignore.case=T), "B1",
ifelse(grepl("ể", f0_reports$syllable.name, ignore.case=T), "C1",
ifelse(grepl("ễ", f0_reports$syllable.name, ignore.case=T), "C2",
ifelse(grepl("ệ", f0_reports$syllable.name, ignore.case=T), "B2",
ifelse(grepl("u", f0_reports$syllable.name, ignore.case=T), "A1",
ifelse(grepl("ù", f0_reports$syllable.name, ignore.case=T), "A2",
ifelse(grepl("ú", f0_reports$syllable.name, ignore.case=T), "B1",
ifelse(grepl("ủ", f0_reports$syllable.name, ignore.case=T), "C1",
ifelse(grepl("ũ", f0_reports$syllable.name, ignore.case=T), "C2",
ifelse(grepl("ụ", f0_reports$syllable.name, ignore.case=T), "B2",
ifelse(grepl("ộ", f0_reports$syllable.name, ignore.case=T), "B2","NA")))))))))))))))))))))))))

# Assigning if the token is single (1) or not (0). Single tokens were produced in isolation.
# Otherwise they were produced in carrier sentences.
f0_reports$single <- ifelse(grepl("single", f0_reports$sound.name), 1, 0)

# Convert categorical variables to factor levels.
f0_reports$gender <- as.factor(f0_reports$gender)
f0_reports$single <- as.factor(f0_reports$single)
f0_reports$tone <- as.factor(f0_reports$tone)
f0_reports$noise <- as.factor(f0_reports$noise)

summary(f0_reports)

```


## Aggregate Tone Analysis

### F0 Plots
```{r}

# Note that I only plot the 2nd to 19th sampled points.
# Extract F0 of all tones A1 (574 instances)
par(mfrow=c(2,4))
A1 <- data.matrix(f0_reports[f0_reports$tone=="A1",-c(1,c(22:26))])
plot(colMeans(A1)[3:18], type="l", xlim=c(1, 18), ylim=c(60, 200))
# # Do the same for the other tones
A2 <- data.matrix(f0_reports[f0_reports$tone=="A2",-c(1,c(22:26))])
plot(colMeans(A2)[3:18], type="l", xlim=c(1, 18), ylim=c(60, 200))

B1 <- data.matrix(f0_reports[f0_reports$tone=="B1",-c(1,c(22:26))])
plot(colMeans(B1)[3:18], type="l", xlim=c(1, 18), ylim=c(60, 200))

B2 <- data.matrix(f0_reports[f0_reports$tone=="B2",-c(1,c(22:26))])
plot(colMeans(B2)[3:18], type="l", xlim=c(1, 18), ylim=c(60, 250))

C1 <- data.matrix(f0_reports[f0_reports$tone=="C1",-c(1,c(22:26))])
plot(colMeans(C1)[3:18], type="l", xlim=c(1, 18), ylim=c(60, 250))

C2 <- data.matrix(f0_reports[f0_reports$tone=="C2",-c(1,c(22:26))])
plot(colMeans(C2)[3:18], type="l", xlim=c(1, 18), ylim=c(60, 250))

D1 <- data.matrix(f0_reports[f0_reports$tone=="D1",-c(1,c(22:26))])
plot(colMeans(D1)[3:18], type="l", xlim=c(1, 18), ylim=c(60, 250))

D2 <- data.matrix(f0_reports[f0_reports$tone=="D2",-c(1,c(22:26))])
plot(colMeans(D2)[3:18], type="l", xlim=c(1, 18), ylim=c(60, 250))
```

### F0 contours on the same plot.
```{r}
plot(colMeans(A1)[3:18], type="b", xlim=c(1, 18), ylim=c(60, 200), xlab="Sampling points", ylab="F0 (Hz)")
lines(colMeans(A2)[3:18],col="green", type="b", pch=19)
lines(colMeans(B1)[3:18],col="red",  type="b", pch=19)
lines(colMeans(B2)[3:18],col="purple", type="b",  pch=19)
lines(colMeans(C1)[3:18],col="blue",  type="b", pch=19)
lines(colMeans(C2)[3:18],col="orange",  type="b", pch=19)

# Add a legend
legend(1, 100, legend=c("A1", "A2", "B1", "B2", "C1", "C2"),
       col=c("black", "green", "red", "purple", "blue", "orange"), lty=1:2, cex=0.7)
```



Plot B1, B2, D1, D2 together.

```{r}
plot(colMeans(B1)[3:18], type="b", col="blue", xlim=c(1, 18), ylim=c(60, 220), , xlab="Sampling points", ylab="F0 (Hz)")
lines(colMeans(B2)[3:18],col="purple", type="b",  pch=19)
lines(colMeans(D1)[3:18],col="pink",  type="b", pch=19)  ## D1 has a very strange contour.
lines(colMeans(D2)[3:18],col="black",  type="b", pch=19)

# Add a legend
legend(1, 100, legend=c("B1", "B2", "D1", "D2"),
       col=c("blue", "orange", "pink", "black"), lty=1:2, cex=0.7)
```

### Estimate Regression Lines for Each Tone on Aggregate.
```{r}

regression_report <- function(tone) {
  lm_tone <- lm(colMeans(tone)[3:18]~ c(1:16))
  plot(colMeans(tone)[3:18], pch = 16, cex = 1.3, xlim=c(1, 18), ylim=c(60, 250))
  abline(lm(colMeans(tone)[3:18] ~ c(1:16)))
  return(lm_tone)
}

```

```{r}
## Call regression_report on all tones.
regression_report(A1)
regression_report(A2)
regression_report(B1)
regression_report(B2)
regression_report(C1)
regression_report(C2)
regression_report(D1)
regression_report(D2)

```
## Tone Contour Analysis on Different Noise Levels (Separately by Single vs. Tokens in Carriers)

```{r}
# Plot F0 contours according to different levels

filter_f0 <- function(tone, noise, s) {
  if (s==TRUE) {
    tone_matrix <- data.matrix(f0_reports[f0_reports$tone==tone & f0_reports$noise==noise & f0_reports$single=="1",-c(1,c(22:26))])
    return(tone_matrix)
  } else {
    tone_matrix <- data.matrix(f0_reports[f0_reports$tone==tone & f0_reports$noise==noise & f0_reports$single=="0",-c(1,c(22:26))])
    return(tone_matrix)
  }
}

plot_f0_by_noise <- function(tone, s) {
  if (s==TRUE) {
    tone_quiet <- data.matrix(f0_reports[f0_reports$tone==tone & f0_reports$noise=="0" & f0_reports$single=="1",-c(1,c(22:26))])
  tone_78 <- data.matrix(f0_reports[f0_reports$tone==tone & f0_reports$noise=="78" & f0_reports$single=="1",-c(1,c(22:26))])
  tone_90 <- data.matrix(f0_reports[f0_reports$tone==tone & f0_reports$noise=="90" & f0_reports$single=="1",-c(1,c(22:26))])
  plot(colMeans(tone_quiet)[3:18], type="l", xlim=c(1, 18), ylim=c(70, 250), xlab="Sampling points", ylab=paste("F0 for single tokens, tone", tone))
  lines(colMeans(tone_78)[3:18],col="blue")
  lines(colMeans(tone_90)[3:18],col="red")
  legend(1, 100, legend=c("quiet", "78dB", "90dB"),
       col=c("black", "blue", "red"), lty=1:2, cex=0.7)
  # Find regression coefficients and return them.
  c(summary(regression_report(tone_quiet))$coefficients[2,1], summary(regression_report(tone_78))$coefficients[2,1], summary(regression_report(tone_90))$coefficients[2,1])
  } else {
    tone_quiet <- data.matrix(f0_reports[f0_reports$tone==tone & f0_reports$noise=="0" & f0_reports$single=="0",-c(1,c(22:26))])
  tone_78 <- data.matrix(f0_reports[f0_reports$tone==tone & f0_reports$noise=="78" & f0_reports$single=="0",-c(1,c(22:26))])
  tone_90 <- data.matrix(f0_reports[f0_reports$tone==tone & f0_reports$noise=="90" & f0_reports$single=="0",-c(1,c(22:26))])
  plot(colMeans(tone_quiet)[3:18], type="l", xlim=c(1, 18), ylim=c(70, 250), xlab="Sampling points", ylab=paste("F0 for tokens in carrier, tone", tone))
  lines(colMeans(tone_78)[3:18],col="blue")
  lines(colMeans(tone_90)[3:18],col="red")
    legend(1, 100, legend=c("quiet", "78dB", "90dB"),
       col=c("black", "blue", "red"), lty=1:2, cex=0.7)
  c(summary(regression_report(tone_quiet))$coefficients[2,1], summary(regression_report(tone_78))$coefficients[2,1], summary(regression_report(tone_90))$coefficients[2,1])
  }
}

coefficient_reports <- function(tone_matrix) {
  lm_model <- lm(colMeans(tone_matrix)[3:18] ~ c(1:16))
  summary(lm_model)$coefficients[2,1]
}
```

### F0 and regression line for tone A1, single tokens.
```{r}
plot_f0_by_noise("A1", TRUE)
```


### F0 and regression line for tone A2, single tokens.
```{r}
plot_f0_by_noise("A2",TRUE)
```


### F0 and regression line for tone B1, single tokens.
```{r}
plot_f0_by_noise("B1", TRUE)
```


### F0 and regression line for tone B2, single tokens.
```{r}
plot_f0_by_noise("B2",TRUE)
```

### F0 and regression line for tone C1, single tokens.
```{r}
plot_f0_by_noise("C1",TRUE)
```

###F0 and regression line for tone C2, single tokens.
```{r}
plot_f0_by_noise("C2", TRUE)
```

### F0 and regression line for tone D1, single tokens.
```{r}
plot_f0_by_noise("D1",TRUE)
```

### F0 and regression line for tone D2, single tokens.
```{r}
plot_f0_by_noise("D2", TRUE)
```

### F0 and regression line for tone A1, tokens in carriers.
```{r}
plot_f0_by_noise("A1", FALSE)
```

### F0 and regression line for tone A2, tokens in carriers.
```{r}
plot_f0_by_noise("A2",FALSE)
```

### F0 and regression line for tone B1, tokens in carriers.
```{r}
plot_f0_by_noise("B1", FALSE)
```

### F0 and regression line for tone B2, tokens in carriers.
```{r}
plot_f0_by_noise("B2",FALSE)
```

### F0 and regression line for tone C1, tokens in carriers.
```{r}
plot_f0_by_noise("C1",FALSE)
```

### F0 and regression line for tone C2, tokens in carriers.
```{r}
plot_f0_by_noise("C2", FALSE)
```

### F0 and regression line for tone D1, tokens in carriers.
```{r}
plot_f0_by_noise("D1",FALSE)
```

### F0 and regression line for tone D2, tokens in carriers.
```{r}
plot_f0_by_noise("D2", FALSE)
```

## All coefficients together in a table.
```{r}
# Table of tone * noise (20x3)
tab <- matrix(,nrow=16, ncol=3, byrow=TRUE)
colnames(tab) <- c('quiet','noise 78','noise 90')
rownames(tab) <- c('A1 single','A2 single','B1 single', 'B2 single',
                   'C1 single','C2 single','D1 single', 'D2 single',
                   'A1 carrier','A2 carrier','B1 carrier', 'B2 carrier',
                   'C1 carrier','C2 carrier','D1 carrier', 'D2 carrier')
tab <- as.table(tab)
tab[1,1] = coefficient_reports(filter_f0("A1", "0", TRUE))
tab[1,2] = coefficient_reports(filter_f0("A1", "78", TRUE))
tab[1,3] = coefficient_reports(filter_f0("A1", "90", TRUE))

tab[2,1] = coefficient_reports(filter_f0("A2", "0", TRUE))
tab[2,2] = coefficient_reports(filter_f0("A2", "78", TRUE))
tab[2,3] = coefficient_reports(filter_f0("A2", "90", TRUE))

tab[3,1] = coefficient_reports(filter_f0("B1", "0", TRUE))
tab[3,2] = coefficient_reports(filter_f0("B1", "78", TRUE))
tab[3,3] = coefficient_reports(filter_f0("B1", "90", TRUE))

tab[4,1] = coefficient_reports(filter_f0("B2", "0", TRUE))
tab[4,2] = coefficient_reports(filter_f0("B2", "78", TRUE))
tab[4,3] = coefficient_reports(filter_f0("B2", "90", TRUE))

tab[5,1] = coefficient_reports(filter_f0("C1", "0", TRUE))
tab[5,2] = coefficient_reports(filter_f0("C1", "78", TRUE))
tab[5,3] = coefficient_reports(filter_f0("C1", "90", TRUE))

tab[6,1] = coefficient_reports(filter_f0("C2", "0", TRUE))
tab[6,2] = coefficient_reports(filter_f0("C2", "78", TRUE))
tab[6,3] = coefficient_reports(filter_f0("C2", "90", TRUE))

tab[7,1] = coefficient_reports(filter_f0("D1", "0", TRUE))
tab[7,2] = coefficient_reports(filter_f0("D1", "78", TRUE))
tab[7,3] = coefficient_reports(filter_f0("D1", "90", TRUE))

tab[8,1] = coefficient_reports(filter_f0("D2", "0", TRUE))
tab[8,2] = coefficient_reports(filter_f0("D2", "78", TRUE))
tab[8,3] = coefficient_reports(filter_f0("D2", "90", TRUE))

# carrier
tab[9,1] = coefficient_reports(filter_f0("A1", "0", FALSE))
tab[9,2] = coefficient_reports(filter_f0("A1", "78", FALSE))
tab[9,3] = coefficient_reports(filter_f0("A1", "90", FALSE))

tab[10,1] = coefficient_reports(filter_f0("A2", "0", FALSE))
tab[10,2] = coefficient_reports(filter_f0("A2", "78", FALSE))
tab[10,3] = coefficient_reports(filter_f0("A2", "90", FALSE))

tab[11,1] = coefficient_reports(filter_f0("B1", "0", FALSE))
tab[11,2] = coefficient_reports(filter_f0("B1", "78", FALSE))
tab[11,3] = coefficient_reports(filter_f0("B1", "90", FALSE))

tab[12,1] = coefficient_reports(filter_f0("B2", "0", FALSE))
tab[12,2] = coefficient_reports(filter_f0("B2", "78", FALSE))
tab[12,3] = coefficient_reports(filter_f0("B2", "90", FALSE))

tab[13,1] = coefficient_reports(filter_f0("C1", "0", FALSE))
tab[13,2] = coefficient_reports(filter_f0("C1", "78", FALSE))
tab[13,3] = coefficient_reports(filter_f0("C1", "90", FALSE))

tab[14,1] = coefficient_reports(filter_f0("C2", "0", FALSE))
tab[14,2] = coefficient_reports(filter_f0("C2", "78", FALSE))
tab[14,3] = coefficient_reports(filter_f0("C2", "90", FALSE))

tab[15,1] = coefficient_reports(filter_f0("D1", "0", FALSE))
tab[15,2] = coefficient_reports(filter_f0("D1", "78", FALSE))
tab[15,3] = coefficient_reports(filter_f0("D1", "90", FALSE))

tab[16,1] = coefficient_reports(filter_f0("D2", "0", FALSE))
tab[16,2] = coefficient_reports(filter_f0("D2", "78", FALSE))
tab[16,3] = coefficient_reports(filter_f0("D2", "90", FALSE))

print(tab)
```

## Tone Contour Analysis on Different Noise Levels on Vowel A.
```{r}

# Add vowel annotation.

f0_reports$vowel <- ifelse(grepl("a", f0_reports$syllable.name, ignore.case=T), "A",
ifelse(grepl("à", f0_reports$syllable.name, ignore.case=T), "A",
ifelse(grepl("á", f0_reports$syllable.name, ignore.case=T), "A",
ifelse(grepl("ả", f0_reports$syllable.name, ignore.case=T), "A",
ifelse(grepl("ã", f0_reports$syllable.name, ignore.case=T), "A",
ifelse(grepl("ạ", f0_reports$syllable.name, ignore.case=T), "A",
ifelse(grepl("ê", f0_reports$syllable.name, ignore.case=T), "E",
ifelse(grepl("ề", f0_reports$syllable.name, ignore.case=T), "E",
ifelse(grepl("ế", f0_reports$syllable.name, ignore.case=T), "E",
ifelse(grepl("ể", f0_reports$syllable.name, ignore.case=T), "E",
ifelse(grepl("ễ", f0_reports$syllable.name, ignore.case=T), "E",
ifelse(grepl("ệ", f0_reports$syllable.name, ignore.case=T), "E",
ifelse(grepl("u", f0_reports$syllable.name, ignore.case=T), "U",
ifelse(grepl("ù", f0_reports$syllable.name, ignore.case=T), "U",
ifelse(grepl("ú", f0_reports$syllable.name, ignore.case=T), "U",
ifelse(grepl("ủ", f0_reports$syllable.name, ignore.case=T), "U",
ifelse(grepl("ũ", f0_reports$syllable.name, ignore.case=T), "U",
ifelse(grepl("ụ", f0_reports$syllable.name, ignore.case=T), "U",
ifelse(grepl("ộ", f0_reports$syllable.name, ignore.case=T), "O","NA")))))))))))))))))))

# Convert vowel types to a factor variable
f0_reports$vowel <- as.factor(f0_reports$vowel)
head(f0_reports, 30)
```



```{r}
# Plot F0 contour by vowels (not distinguishing single tokens or tokens in carriers.)

plot_f0_by_vowels <- function(tone, vowel) {
  if (vowel=="A") {
    tone_quiet <- data.matrix(f0_reports[f0_reports$tone==tone & f0_reports$noise=="0" & f0_reports$vowel=="A",-c(1,c(22:26))])
  tone_78 <- data.matrix(f0_reports[f0_reports$tone==tone & f0_reports$noise=="78" & f0_reports$vowel=="A",-c(1,c(22:26))])
  tone_90 <- data.matrix(f0_reports[f0_reports$tone==tone & f0_reports$noise=="90" & f0_reports$vowel=="A",-c(1,c(22:26))])
  plot(colMeans(tone_quiet)[3:18], type="l", xlim=c(1, 18), ylim=c(70, 250), xlab="Sampling points", ylab=paste("F0 for vowel A, tone", tone))
  lines(colMeans(tone_78)[3:18],col="blue")
  lines(colMeans(tone_90)[3:18],col="red")
  # Find regression coefficients and return them.
  c(summary(regression_report(tone_quiet))$coefficients[2,1], summary(regression_report(tone_78))$coefficients[2,1], summary(regression_report(tone_90))$coefficients[2,1])
  } else if (vowel=="E") {
    tone_quiet <- data.matrix(f0_reports[f0_reports$tone==tone & f0_reports$noise=="0" & f0_reports$vowel=="E",-c(1,c(22:26))])
  tone_78 <- data.matrix(f0_reports[f0_reports$tone==tone & f0_reports$noise=="78" & f0_reports$vowel=="E",-c(1,c(22:26))])
  tone_90 <- data.matrix(f0_reports[f0_reports$tone==tone & f0_reports$noise=="90" & f0_reports$vowel=="E",-c(1,c(22:26))])
  plot(colMeans(tone_quiet)[3:18], type="l", xlim=c(1, 18), ylim=c(70, 250), xlab="Sampling points", ylab=paste("F0 for vowel E, tone", tone))
  lines(colMeans(tone_78)[3:18],col="blue")
  lines(colMeans(tone_90)[3:18],col="red")
  # Find regression coefficients and return them.
  c(summary(regression_report(tone_quiet))$coefficients[2,1], summary(regression_report(tone_78))$coefficients[2,1], summary(regression_report(tone_90))$coefficients[2,1])
  } else {
  tone_quiet <- data.matrix(f0_reports[f0_reports$tone==tone & f0_reports$noise=="0" & f0_reports$vowel=="U",-c(1,c(22:26))])
  tone_78 <- data.matrix(f0_reports[f0_reports$tone==tone & f0_reports$noise=="78" & f0_reports$vowel=="U",-c(1,c(22:26))])
  tone_90 <- data.matrix(f0_reports[f0_reports$tone==tone & f0_reports$noise=="90" & f0_reports$vowel=="U",-c(1,c(22:26))])
  plot(colMeans(tone_quiet)[3:18], type="l", xlim=c(1, 18), ylim=c(70, 250), xlab="Sampling points", ylab=paste("F0 for vowel U, tone", tone))
  lines(colMeans(tone_78)[3:18],col="blue")
  lines(colMeans(tone_90)[3:18],col="red")
  # Find regression coefficients and return them.
  c(summary(regression_report(tone_quiet))$coefficients[2,1], summary(regression_report(tone_78))$coefficients[2,1], summary(regression_report(tone_90))$coefficients[2,1])
  }
}

# Call functions

```


## Tone A1 on A
```{r}
plot_f0_by_vowels("A1", "A")
```

## Tone A1 on E
```{r}
plot_f0_by_vowels("A1", "E")
```

## Tone A1 on U
```{r}
plot_f0_by_vowels("A1", "U")
```

## Tone A2 on A
```{r}
plot_f0_by_vowels("A2", "A")
```

## Tone A2 on E
```{r}
plot_f0_by_vowels("A2", "E")
```

## Tone A2 on U
```{r}
plot_f0_by_vowels("A2", "U")
```


## Tone B1 on A
```{r}
plot_f0_by_vowels("B1", "A")
```

## Tone B1 on E
```{r}
plot_f0_by_vowels("B1", "E")
```

## Tone B1 on U
```{r}
plot_f0_by_vowels("B1", "U")
```

## Tone B2 on A
```{r}
plot_f0_by_vowels("B2", "A")
```

## Tone B2 on E
```{r}
plot_f0_by_vowels("B2", "E")
```

## Tone B2 on U
```{r}
plot_f0_by_vowels("B2", "U")
```

## Tone C1 on A
```{r}
plot_f0_by_vowels("C1", "A")
```

## Tone C1 on E
```{r}
plot_f0_by_vowels("C1", "E")
```

## Tone C1 on U
```{r}
plot_f0_by_vowels("C1", "U")
```

## Tone C2 on A
```{r}
plot_f0_by_vowels("C2", "A")
```

## Tone C2 on E
```{r}
plot_f0_by_vowels("C2", "E")
```

## Tone C2 on U
```{r}
plot_f0_by_vowels("C2", "U")
```


## Tone D1 on A
```{r}
plot_f0_by_vowels("D1", "A")
```

## Tone D1 on E
```{r}
plot_f0_by_vowels("D1", "E")
```

## Tone D1 on U
```{r}
plot_f0_by_vowels("D1", "U")
```

## Tone D2 on A
```{r}
plot_f0_by_vowels("D2", "A")
```

## Tone D2 on E
```{r}
plot_f0_by_vowels("D2", "E")
```

## Tone D2 on U
```{r}
plot_f0_by_vowels("D2", "U")
```

### Slope coefficients in a table
# Plot F0 contours according to different levels

```{r}
filter_f0 <- function(tone, vowel, noise) {
    tone_matrix <- data.matrix(f0_reports[f0_reports$tone==tone & f0_reports$noise==noise & f0_reports$vowel==vowel,-c(1,c(22:26))])
    return(tone_matrix)
}

# Table of tone * noise (24x3)
tab_vowel <- matrix(,nrow=24, ncol=3, byrow=TRUE)
colnames(tab_vowel) <- c('quiet','noise 78','noise 90')
rownames(tab_vowel) <- c('A1 A','A2 A','B1 A', 'B2 A',
                   'C1 A','C2 A','D1 A', 'D2 A',
                   'A1 E','A2 E','B1 E', 'B2 E',
                   'C1 E','C2 E','D1 E', 'D2 E',
                   'A1 U','A2 U','B1 U', 'B2 U',
                   'C1 U','C2 U','D1 U', 'D2 U')
tab_vowel <- as.table(tab_vowel)
tab_vowel[1,1] = coefficient_reports(filter_f0("A1", "A", "0"))
tab_vowel[1,2] = coefficient_reports(filter_f0("A1", "A", "78"))
tab_vowel[1,3] = coefficient_reports(filter_f0("A1", "A", "90"))

tab_vowel[2,1] = coefficient_reports(filter_f0("A2", "A", "0"))
tab_vowel[2,2] = coefficient_reports(filter_f0("A2", "A", "78"))
tab_vowel[2,3] = coefficient_reports(filter_f0("A2", "A", "90"))

tab_vowel[3,1] = coefficient_reports(filter_f0("B1", "A", "0"))
tab_vowel[3,2] = coefficient_reports(filter_f0("B1", "A", "78"))
tab_vowel[3,3] = coefficient_reports(filter_f0("B1", "A", "90"))

tab_vowel[4,1] = coefficient_reports(filter_f0("B2", "A", "0"))
tab_vowel[4,2] = coefficient_reports(filter_f0("B2", "A", "78"))
tab_vowel[4,3] = coefficient_reports(filter_f0("B2", "A", "90"))

tab_vowel[5,1] = coefficient_reports(filter_f0("C1", "A", "0"))
tab_vowel[5,2] = coefficient_reports(filter_f0("C1", "A", "78"))
tab_vowel[5,3] = coefficient_reports(filter_f0("C1", "A", "90"))

tab_vowel[6,1] = coefficient_reports(filter_f0("C2", "A", "0"))
tab_vowel[6,2] = coefficient_reports(filter_f0("C2", "A", "78"))
tab_vowel[6,3] = coefficient_reports(filter_f0("C2", "A", "90"))

tab_vowel[7,1] = coefficient_reports(filter_f0("D1", "A", "0"))
tab_vowel[7,2] = coefficient_reports(filter_f0("D1", "A", "78"))
tab_vowel[7,3] = coefficient_reports(filter_f0("D1", "A", "90"))

tab_vowel[8,1] = coefficient_reports(filter_f0("D2", "A", "0"))
tab_vowel[8,2] = coefficient_reports(filter_f0("D2", "A", "78"))
tab_vowel[8,3] = coefficient_reports(filter_f0("D2", "A", "90"))

tab_vowel[9,1] = coefficient_reports(filter_f0("A1", "E", "0"))
tab_vowel[9,2] = coefficient_reports(filter_f0("A1", "E", "78"))
tab_vowel[9,3] = coefficient_reports(filter_f0("A1", "E", "90"))

tab_vowel[10,1] = coefficient_reports(filter_f0("A2", "E", "0"))
tab_vowel[10,2] = coefficient_reports(filter_f0("A2", "E", "78"))
tab_vowel[10,3] = coefficient_reports(filter_f0("A2", "E", "90"))

tab_vowel[11,1] = coefficient_reports(filter_f0("B1", "E", "0"))
tab_vowel[11,2] = coefficient_reports(filter_f0("B1", "E", "78"))
tab_vowel[11,3] = coefficient_reports(filter_f0("B1", "E", "90"))

tab_vowel[12,1] = coefficient_reports(filter_f0("B2", "E", "0"))
tab_vowel[12,2] = coefficient_reports(filter_f0("B2", "E", "78"))
tab_vowel[12,3] = coefficient_reports(filter_f0("B2", "E", "90"))

tab_vowel[13,1] = coefficient_reports(filter_f0("C1", "E", "0"))
tab_vowel[13,2] = coefficient_reports(filter_f0("C1", "E", "78"))
tab_vowel[13,3] = coefficient_reports(filter_f0("C1", "E", "90"))

tab_vowel[14,1] = coefficient_reports(filter_f0("C2", "E", "0"))
tab_vowel[14,2] = coefficient_reports(filter_f0("C2", "E", "78"))
tab_vowel[14,3] = coefficient_reports(filter_f0("C2", "E", "90"))

tab_vowel[15,1] = coefficient_reports(filter_f0("D1", "E", "0"))
tab_vowel[15,2] = coefficient_reports(filter_f0("D1", "E", "78"))
tab_vowel[15,3] = coefficient_reports(filter_f0("D1", "E", "90"))

tab_vowel[16,1] = coefficient_reports(filter_f0("D2", "E", "0"))
tab_vowel[16,2] = coefficient_reports(filter_f0("D2", "E", "78"))
tab_vowel[16,3] = coefficient_reports(filter_f0("D2", "E", "90"))

tab_vowel[17,1] = coefficient_reports(filter_f0("A1", "U", "0"))
tab_vowel[17,2] = coefficient_reports(filter_f0("A1", "U", "78"))
tab_vowel[17,3] = coefficient_reports(filter_f0("A1", "U", "90"))

tab_vowel[18,1] = coefficient_reports(filter_f0("A2", "U", "0"))
tab_vowel[18,2] = coefficient_reports(filter_f0("A2", "U", "78"))
tab_vowel[18,3] = coefficient_reports(filter_f0("A2", "U", "90"))

tab_vowel[19,1] = coefficient_reports(filter_f0("B1", "U", "0"))
tab_vowel[19,2] = coefficient_reports(filter_f0("B1", "U", "78"))
tab_vowel[19,3] = coefficient_reports(filter_f0("B1", "U", "90"))

tab_vowel[20,1] = coefficient_reports(filter_f0("B2", "U", "0"))
tab_vowel[20,2] = coefficient_reports(filter_f0("B2", "U", "78"))
tab_vowel[20,3] = coefficient_reports(filter_f0("B2", "U", "90"))

tab_vowel[21,1] = coefficient_reports(filter_f0("C1", "U", "0"))
tab_vowel[21,2] = coefficient_reports(filter_f0("C1", "U", "78"))
tab_vowel[21,3] = coefficient_reports(filter_f0("C1", "U", "90"))

tab_vowel[22,1] = coefficient_reports(filter_f0("C2", "U", "0"))
tab_vowel[22,2] = coefficient_reports(filter_f0("C2", "U", "78"))
tab_vowel[22,3] = coefficient_reports(filter_f0("C2", "U", "90"))

tab_vowel[23,1] = coefficient_reports(filter_f0("D1", "U", "0"))
tab_vowel[23,2] = coefficient_reports(filter_f0("D1", "U", "78"))
tab_vowel[23,3] = coefficient_reports(filter_f0("D1", "U", "90"))

tab_vowel[24,1] = coefficient_reports(filter_f0("D2", "U", "0"))
tab_vowel[24,2] = coefficient_reports(filter_f0("D2", "U", "78"))
tab_vowel[24,3] = coefficient_reports(filter_f0("D2", "U", "90"))


tab_vowel


```