---
title: "Tone Classification"
author: "Giang Le"
date: "10/28/2021"
output:
  html_document:
    df_print: paged
  pdf_document: 
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Import and Preprocessing
Our data contains 15 voice reports from 15 recording sessions.

```{r}
# Read in all the voice reports.
dataFiles <- lapply(Sys.glob("*/channel1/acoustic_measurements_unique_*.csv"), read.csv)
```

```{r}
## Add the following categorical predictors.
# Gender: F and M (done)
# Noise type: quiet, 78 or 90 (done)
# Single token or token in a sentence
# Syllable type 
# Tone

# Converting to DataFrames

f_1_78 <- as.data.frame(dataFiles[1])
f_1_90 <- as.data.frame(dataFiles[2])
f_1_q <- as.data.frame(dataFiles[3])

f_2_78 <- as.data.frame(dataFiles[4])
f_2_90 <- as.data.frame(dataFiles[5])
f_2_q <- as.data.frame(dataFiles[6])

m_1_78 <- as.data.frame(dataFiles[7])
m_1_90 <- as.data.frame(dataFiles[8])
m_1_q <- as.data.frame(dataFiles[9])

m_2_78 <- as.data.frame(dataFiles[10])
m_2_90 <- as.data.frame(dataFiles[11])
m_2_q <- as.data.frame(dataFiles[12])

m_3_78 <- as.data.frame(dataFiles[13])
m_3_90 <- as.data.frame(dataFiles[14])
m_3_q <- as.data.frame(dataFiles[15])

```

```{r}
# Assigning gender variable (0 for female and 1 for male)
f_1_78$gender = 0
f_1_90$gender = 0
f_1_q$gender = 0

f_2_78$gender = 0
f_2_90$gender = 0
f_2_q$gender = 0

m_1_78$gender = 1
m_1_90$gender = 1
m_1_q$gender = 1

m_2_78$gender = 1
m_2_90$gender = 1
m_2_q$gender = 1

m_3_78$gender = 1
m_3_90$gender = 1
m_3_q$gender = 1

# Assigning noise level
f_1_78$noise = 78
f_1_90$noise = 90
f_1_q$noise = 0


f_2_78$noise = 78
f_2_90$noise = 90
f_2_q$noise = 0

m_1_78$noise = 78
m_1_90$noise = 90
m_1_q$noise = 0

m_2_78$noise = 78
m_2_90$noise = 90
m_2_q$noise = 0

m_3_78$noise = 78
m_3_90$noise = 90
m_3_q$noise = 0

# Assigning speaker code
f_1_78$speaker = "f-1"
f_1_90$speaker = "f-1"
f_1_q$speaker = "f-1"

f_2_78$speaker = "f-2"
f_2_90$speaker = "f-2"
f_2_q$speaker = "f-2"

m_1_78$speaker = "m-1"
m_1_90$speaker = "m-1"
m_1_q$speaker = "m-1"

m_2_78$speaker = "m-2"
m_2_90$speaker = "m-2"
m_2_q$speaker = "m-2"

m_3_78$speaker = "m-3"
m_3_90$speaker = "m-3"
m_3_q$speaker = "m-3"
```

```{r}
### Concatenate all dataframes
voice_reports <- rbind(f_1_78, f_1_90, f_1_q,
                       f_2_78, f_2_90, f_2_q,
                       m_1_78, m_1_90, m_1_q,
                       m_2_78,m_2_90, m_2_q,
                       m_3_78, m_3_90, m_3_q)



## Drop intervals that don't matter
voice_reports <- voice_reports[!(endsWith(voice_reports$sound.name,"_")),]
dim(voice_reports)
head(voice_reports,10)
```

```{r}
# Assigning if the token is single (1) or not (0).
voice_reports$single <- ifelse(grepl("single", voice_reports$sound.name), 1, 0)

# Assign syllable shapes (do later)


# Assign tone values
# voice_reports$tone <- ifelse(grepl("a", voice_reports$sound.name, ignore.case=T), "A1",
#                       ifelse(grepl("à", voice_reports$sound.name, ignore.case=T), "A2",
#                       ifelse(grepl("á", voice_reports$sound.name, ignore.case=T), "B1",
#                       ifelse(grepl("ả", voice_reports$sound.name, ignore.case=T), "C1",
#                       ifelse(grepl("ã", voice_reports$sound.name, ignore.case=T), "C2",
#                       ifelse(grepl("ạ", voice_reports$sound.name, ignore.case=T), "B2",
#                       ifelse(grepl("ê", voice_reports$sound.name, ignore.case=T), "A1",
#                       ifelse(grepl("ề", voice_reports$sound.name, ignore.case=T), "A2",
#                       ifelse(grepl("ế", voice_reports$sound.name, ignore.case=T), "B1",
#                       ifelse(grepl("ể", voice_reports$sound.name, ignore.case=T), "C1",
#                       ifelse(grepl("ễ", voice_reports$sound.name, ignore.case=T), "C2",
#                       ifelse(grepl("ệ", voice_reports$sound.name, ignore.case=T), "B2",
#                       ifelse(grepl("u", voice_reports$sound.name, ignore.case=T), "A1",
#                       ifelse(grepl("ù", voice_reports$sound.name, ignore.case=T), "A2",
#                       ifelse(grepl("ú", voice_reports$sound.name, ignore.case=T), "B1",
#                       ifelse(grepl("ủ", voice_reports$sound.name, ignore.case=T), "C1",
#                       ifelse(grepl("ũ", voice_reports$sound.name, ignore.case=T), "C2",
#                       ifelse(grepl("ụ", voice_reports$sound.name, ignore.case=T), "B2",
#                       ifelse(grepl("ộ", voice_reports$sound.name, ignore.case=T), "B2","NA")))))))))))))))))))

voice_reports$tone <- ifelse(grepl("a", voice_reports$sound.name, ignore.case=T), "A1",
ifelse(grepl("_tát", voice_reports$sound.name, ignore.case=T), "D1",
ifelse(grepl("_tạt", voice_reports$sound.name, ignore.case=T), "D2",
ifelse(grepl("_tết", voice_reports$sound.name, ignore.case=T), "D1",
ifelse(grepl("_tệt", voice_reports$sound.name, ignore.case=T), "D2",
ifelse(grepl("_tút", voice_reports$sound.name, ignore.case=T), "D1",
ifelse(grepl("_tụt", voice_reports$sound.name, ignore.case=T), "D2",
ifelse(grepl("à", voice_reports$sound.name, ignore.case=T), "A2",
ifelse(grepl("á", voice_reports$sound.name, ignore.case=T), "B1",
ifelse(grepl("ả", voice_reports$sound.name, ignore.case=T), "C1",
ifelse(grepl("ã", voice_reports$sound.name, ignore.case=T), "C2",
ifelse(grepl("ạ", voice_reports$sound.name, ignore.case=T), "B2",
ifelse(grepl("ê", voice_reports$sound.name, ignore.case=T), "A1",
ifelse(grepl("ề", voice_reports$sound.name, ignore.case=T), "A2",
ifelse(grepl("ế", voice_reports$sound.name, ignore.case=T), "B1",
ifelse(grepl("ể", voice_reports$sound.name, ignore.case=T), "C1",
ifelse(grepl("ễ", voice_reports$sound.name, ignore.case=T), "C2",
ifelse(grepl("ệ", voice_reports$sound.name, ignore.case=T), "B2",
ifelse(grepl("u", voice_reports$sound.name, ignore.case=T), "A1",
ifelse(grepl("ù", voice_reports$sound.name, ignore.case=T), "A2",
ifelse(grepl("ú", voice_reports$sound.name, ignore.case=T), "B1",
ifelse(grepl("ủ", voice_reports$sound.name, ignore.case=T), "C1",
ifelse(grepl("ũ", voice_reports$sound.name, ignore.case=T), "C2",
ifelse(grepl("ụ", voice_reports$sound.name, ignore.case=T), "B2",
ifelse(grepl("ộ", voice_reports$sound.name, ignore.case=T), "B2","NA")))))))))))))))))))))))))

# Assign phonation types
voice_reports$phonation <- ifelse(grepl("A1", voice_reports$tone, ignore.case=T), "modal",
                      ifelse(grepl("A2", voice_reports$tone, ignore.case=T), "breathy",
                      ifelse(grepl("B1", voice_reports$tone, ignore.case=T), "modal",
                      ifelse(grepl("B2", voice_reports$tone, ignore.case=T), "creaky",
                      ifelse(grepl("C1", voice_reports$tone, ignore.case=T), "creaky",
                      ifelse(grepl("C2", voice_reports$tone, ignore.case=T), "creaky","NA"))))))


# Assign creakiness or not
voice_reports$creaky <- ifelse(grepl("creaky", voice_reports$phonation, ignore.case=T), 1, 0)
voice_reports$creaky <- as.factor(voice_reports$creaky)
head(voice_reports, 20)
```

### Checking
```{r}
# How many values are of each category
length(voice_reports$tone[voice_reports$tone == "A1"])
## [1] 574
length(voice_reports$tone[voice_reports$tone == "A2"])
## [1] 575
length(voice_reports$tone[voice_reports$tone == "B1"])
## [1] 719
length(voice_reports$tone[voice_reports$tone == "B2"])
## [1] 768
length(voice_reports$tone[voice_reports$tone == "C1"])
## [1] 575
length(voice_reports$tone[voice_reports$tone == "C2"])
## [1] 575

length(voice_reports$tone[voice_reports$tone == "D1"])
## [1] 575
length(voice_reports$tone[voice_reports$tone == "D2"])
## [1] 575
length(voice_reports$tone[voice_reports$tone == "NA"])
## [1] 0
```

### Convert categorical values to factors
```{r}
## Not sure if this is necessary for variables already binarily coded.
voice_reports$gender <- as.factor(voice_reports$gender)
voice_reports$noise <- as.factor(voice_reports$noise)
voice_reports$tone <- as.factor(voice_reports$tone)
voice_reports$single <- as.factor(voice_reports$single)
voice_reports$phonation <- as.factor(voice_reports$phonation)
voice_reports$creaky <- as.factor(voice_reports$creaky)
voice_reports$speaker <- as.factor(voice_reports$speaker)

```


### Summary of current data
```{r}
summary(voice_reports)
```

### Clean up undefined values to prepare for Classification
```{r}

## Method 1: Simply drop values that are undefined in jitter and shimmer variables
voice_reports_clean <- voice_reports[!(voice_reports$jitter.local==" --undefined-- " | voice_reports$shimmer.local==" --undefined-- "),]

# Convert two variables to numeric
voice_reports_clean$jitter.local <- as.numeric(voice_reports_clean$jitter.local)
voice_reports_clean$shimmer.local <- as.numeric(voice_reports_clean$shimmer.local)
voice_reports_clean$median.F0 <- as.numeric(voice_reports_clean$median.F0)
voice_reports_clean$mean.F0 <- as.numeric(voice_reports_clean$mean.F0)
voice_reports_clean$sd.F0 <- as.numeric(voice_reports_clean$sd.F0)
voice_reports_clean$min.F0 <- as.numeric(voice_reports_clean$min.F0)
voice_reports_clean$max.F0 <- as.numeric(voice_reports_clean$max.F0)
voice_reports_clean$number.pulses<- as.numeric(voice_reports_clean$number.pulses)
voice_reports_clean$number.periods <- as.numeric(voice_reports_clean$number.periods)
voice_reports_clean$mean.periods <- as.numeric(voice_reports_clean$mean.periods)
#voice_reports_clean$sd.periods <- as.numeric(voice_reports_clean$sd.periods)
voice_reports_clean$jitter.local.abs <- as.numeric(voice_reports_clean$jitter.local.abs)
voice_reports_clean$jitter.rap <- as.numeric(voice_reports_clean$jitter.rap)
voice_reports_clean$jitter.ppq5 <- as.numeric(voice_reports_clean$jitter.ppq5)
voice_reports_clean$shimmer.local.db <- as.numeric(voice_reports_clean$shimmer.local.db)
voice_reports_clean$shimmer.apq3 <- as.numeric(voice_reports_clean$shimmer.apq3)
voice_reports_clean$shimmer.apq5 <- as.numeric(voice_reports_clean$shimmer.apq5)
voice_reports_clean$shimmer.apq11 <- as.numeric(voice_reports_clean$shimmer.apq11)
voice_reports_clean$mean.autocorr <- as.numeric(voice_reports_clean$mean.autocorr)
voice_reports_clean$mean.NHR <- as.numeric(voice_reports_clean$mean.NHR)
voice_reports_clean$mean.HNR <- as.numeric(voice_reports_clean$mean.HNR)
voice_reports_clean$tone <- as.factor(voice_reports_clean$tone)
summary(voice_reports_clean)
dim(voice_reports_clean)
```

### Logistic Regression on Gender
```{r}
logit_gender = glm(gender ~ mean.F0 + total.duration + intensity +  mean.HNR, family = "binomial", data=voice_reports_clean)
summary(logit_gender)
```

### Logistic Regression on Creaky
```{r}
logit_creaky = glm(creaky ~  mean.F0 + total.duration + intensity + spectraltilt + number.pulses + mean.HNR + jitter.local + shimmer.local, family = "binomial", data=voice_reports_clean)
summary(logit_creaky)
```

### Regression with Interaction
```{r}
lm_creaky = lm(jitter.local ~ mean.HNR + noise + mean.HNR*noise, data = voice_reports_clean)
summary(lm_creaky)

```
### Multinomial Regression to predict the Noise Level.
```{r}
# Use the multinom function from the nnet package (Ref: https://stats.idre.ucla.edu/r/dae/multinomial-logistic-regression/)

library("nnet")
# Use the 78 noise level as the reference level
voice_reports_clean$noise2 <- relevel(voice_reports_clean$noise, ref = "78")
multinom_noise <- multinom(noise2 ~ mean.F0 + total.duration + intensity + spectraltilt, data=voice_reports_clean)
summary(multinom_noise)

# The result in general supports our predictions regarding the relationship
#between relative noise levels
# and F0, duration, intensity, etc.

# For instance,
# A one-unit increase in mean F0 is associated with the decrease in the
#log odds of quiet vs. 78 noise level in the amount of 0.0133
# A one-unit increase in mean F0 is associated with the increase in the
#log odds of 90 noise vs. 78 noise in the amount of 0.006

# A one-unit increase in duration is associated with the decrease in the
#log odds of quiet vs. 78 noise level in the amount of 5.795
# A one-unit increase in duration is associated with the increase in the
#log odds of 90 noise vs. 78 noise in the amount of 2.80

# A one-unit increase in intensity is associated with the decrease in the
#log odds of quiet vs. 78 noise level in the amount of 0.35
# A one-unit increase in intensity is associated with the increase in the
#*log odds of 90 noise vs. 78 noise in the amount of 0.24

## Giang to double check this result
# A one-unit increase in spectraltilt is associated with the decrease in the
#log odds of quiet vs. 78 noise level in the amount of 0.066
# A one-unit increase in spectraltilt is associated with the increase in the
#log odds of 90 noise vs. 78 noise in the amount of 0.011
```


## Classification using SMV (ref https://medium.com/@ODSC/build-a-multi-class-support-vector-machine-in-r-abcdd4b7dab6)

```{r}

library(e1071)   
set.seed(777)
n <- nrow(voice_reports_clean)
ntrain <- round(n*0.75)  # 75% for training set
tindex <- sample(n, ntrain)
# Do not include noise predictor in this model yet.
train <- voice_reports_clean[tindex,c("total.duration", "intensity", 
                                      "spectraltilt", "mean.F0", "jitter.local",
                                      "shimmer.local", "mean.HNR", "gender",
                                      "tone")]
test <- voice_reports_clean[-tindex,c("total.duration", "intensity",
                                      "spectraltilt", "mean.F0", "jitter.local",
                                      "shimmer.local", "mean.HNR", "gender",
                                        "tone")]

# Some factors cause any error probably due to not having the same levels between train and test?
svm_model <- svm(tone ~ total.duration + intensity + spectraltilt + mean.F0 + jitter.local
                 + shimmer.local + mean.HNR + gender, data=train, 
          method="C-classification", kernal="radial", 
          gamma=0.1, cost=10)

summary(svm_model)
prediction <- predict(svm_model, test)
confusion <- table(test$tone, prediction)
confusion

# Accuracy
sum(diag(confusion))/sum(confusion) 
```

## Classification using k-means clustering

```{r}
install.packages("caret", repos = "http://cran.us.r-project.org")
library("caret")

set.seed(777)
n <- nrow(voice_reports_clean)
ntrain <- round(n*0.75)  # 75% for training set
tindex <- sample(n, ntrain)
# Do not include noise predictor in this model yet.
train <- voice_reports_clean[tindex,c("total.duration", "intensity",
                                      "spectraltilt", "mean.F0", "jitter.local",
                                      "shimmer.local", "mean.HNR", "gender", "tone")]
test <- voice_reports_clean[-tindex,c("total.duration", "intensity",
                                      "spectraltilt", "mean.F0", "jitter.local",
                                          "shimmer.local", "mean.HNR", "gender", "tone")]
summary(train)
```

```{r}
# Use repeated 5-fold cross-validation, with 3 repeats.
install.packages("kknn", repos = "http://cran.us.r-project.org")
library("kknn")

```

```{r}

# # I set the parameters of trainControl to be method ~ repeatedcv, 5 folds, # and 2 repeats.
# control <- trainControl(method = "repeatedcv", number = 5, repeats=2)
# # Here I train the model using knn, and k values from 1 to 10.
# knn.cvfit <- train(tone ~ ., method = "knn", data = train,
# tuneGrid = data.frame(k = seq(1, 6, 1)), trControl = control)
# 
# plot(knn.cvfit$results$k, 1-knn.cvfit$results$Accuracy,
# xlab = "K", ylab = "Classification Error", type = "b", pch = 19, col = "darkorange")
```

## Vowels plotting 

```{r}
#http://lingtools.uoregon.edu/norm/about_norm1.php

install.packages("vowels", repos='http://cran.us.r-project.org')
library(vowels)
```

```{r}
# Prepare vowels data
vowels <- voice_reports_clean[, c(37, 1, 36, 31, 32, 33)]
vowels$gl.F1 <- NA
vowels$gl.F2 <- NA
vowels$gl.F3 <- NA
# Extracting a substring that contains only the syllable names.
vowels$sound.name <- sapply(strsplit(vowels[,2], split="_", fixed=TRUE), "[", 2)

# Add vowel annotation.
vowels$vowel <- ifelse(grepl("a", vowels$sound.name, ignore.case=T), "a",
ifelse(grepl("à", vowels$sound.name, ignore.case=T), "a",
ifelse(grepl("á", vowels$sound.name, ignore.case=T), "a",
ifelse(grepl("ả", vowels$sound.name, ignore.case=T), "a",
ifelse(grepl("ã", vowels$sound.name, ignore.case=T), "a",
ifelse(grepl("ạ", vowels$sound.name, ignore.case=T), "a",
ifelse(grepl("ê", vowels$sound.name, ignore.case=T), "e",
ifelse(grepl("ề", vowels$sound.name, ignore.case=T), "e",
ifelse(grepl("ế", vowels$sound.name, ignore.case=T), "e",
ifelse(grepl("ể", vowels$sound.name, ignore.case=T), "e",
ifelse(grepl("ễ", vowels$sound.name, ignore.case=T), "e",
ifelse(grepl("ệ", vowels$sound.name, ignore.case=T), "e",
ifelse(grepl("u", vowels$sound.name, ignore.case=T), "u",
ifelse(grepl("ù", vowels$sound.name, ignore.case=T), "u",
ifelse(grepl("ú", vowels$sound.name, ignore.case=T), "u",
ifelse(grepl("ủ", vowels$sound.name, ignore.case=T), "u",
ifelse(grepl("ũ", vowels$sound.name, ignore.case=T), "u",
ifelse(grepl("ụ", vowels$sound.name, ignore.case=T), "u",
ifelse(grepl("ộ", vowels$sound.name, ignore.case=T), "o","NA")))))))))))))))))))

# Convert vowel types to a factor variable
vowels$vowel <- as.factor(vowels$vowel)
vowels$noise <- as.factor(vowels$noise)
vowels <- vowels[,c("speaker", "vowel", "noise", "F1", "F2", "F3", "gl.F1", "gl.F2", "gl.F3")]

# plot only sub-dataframes
vowels_plotting <- function(datamat, noise, speaker) {
  if (speaker == "all") {
    vowels <- datamat[datamat$noise==noise,]
    vowelplot(norm.bark(vowels), title=paste("F1-F2 vowel space for", speaker, "in", noise), color="vowels", labels="noise", color.choice=rainbow(length(unique(vowels[,2]))))
  } else {
    vowels <- datamat[datamat$noise==noise & datamat$speaker==speaker,]
  vowelplot(norm.bark(vowels), title=paste("F1-F2 vowel space for speaker", speaker, "in", noise), color="vowels", labels="noise", color.choice=rainbow(length(unique(vowels[,2]))))
  }
  
}

vowels_plotting(vowels, 0, "f-1")
vowels_plotting(vowels, 78, "f-1")
vowels_plotting(vowels, 90, "f-1")

vowels_plotting(vowels, 0, "f-2")
vowels_plotting(vowels, 78, "f-2")
vowels_plotting(vowels, 90, "f-2")


vowels_plotting(vowels, 0, "m-1")
vowels_plotting(vowels, 78, "m-1")
vowels_plotting(vowels, 90, "m-1")

vowels_plotting(vowels, 0, "m-2")
vowels_plotting(vowels, 78, "m-2")
vowels_plotting(vowels, 90, "m-2")

vowels_plotting(vowels, 0, "m-3")
vowels_plotting(vowels, 78, "m-3")
vowels_plotting(vowels, 90, "m-3")

# par(mfrow=c(2,1))
# vowelplot(compute.means(vowels), shape="vowels")
# vowelplot(compute.means(norm.lobanov(vowels)), shape="vowels")

# par(mfrow=c(1,1))
# g09.means <- compute.means(vowels, speaker="f-1")
# vowelplot(g09.means, color="vowels", labels="none")
# add.spread.vowelplot(vowels, speaker="f-1", sd.mult=1, color="vowels", labels="none")
# # can add annotations to the vowel plots as any other R graph, eg:
# legend("top", legend="Can you guess which vowel is 'BOY'?", col='lightslategrey', bty="n")

```

```{r}
vowels_plotting(vowels, 0, "all")
vowels_plotting(vowels, 78, "all")
vowels_plotting(vowels, 90, "all")
```

## Use phonR to calculate the hull area. Decreasing area found.
```{r}
install.packages("phonR", repos = 'http://cran.us.r-project.org')
library(phonR)

```

```{r}

convexHullArea(vowels$F1, vowels$F2, group=vowels$speaker)

# reduced hull.area
hull.area <- with(vowels, convexHullArea(F1, F2, group=noise))
hull.area
```